---
title: "CRBLM RNAseq Differential Expression"
output:
  pdf_document: default
  html_notebook: default
---
--------

-----------
#installing bioconductor packages
```{r include=FALSE}
#install.packages("devtools")
#install.packages("BiocManager")
#BiocManager::install("TxDb.Rnorvegicus.UCSC.rn6.refGene")
#BiocManager::install("Rsubread")
#BiocManager::install("DESeq2")
#BiocManager::install("Rattus.norvegicus")
#BiocManager::install("EnhancedVolcano")
#BiocManager::install("topGO")
#BiocManager::install("org.Rn.eg.db")
#BiocManager::install("DOSE")
#BiocManager::install("clusterProfiler")
#BiocManager::install("pathview")
#BiocManager::install("fgsea")
```


#Load libraries
```{r include=FALSE}
library(ggplot2)
library(rgl)
library(tidyr)
library(tidyverse)
library(RColorBrewer)
library(expss)
library(dplyr)
library(BiocManager)
library(Rsamtools)
library(GenomicAlignments)
library(BiocParallel)
library(GenomicFeatures)
library(Rsubread)
library(DESeq2)
library(Rsamtools)
library(pheatmap)
library(AnnotationDbi)
library(Rattus.norvegicus)
library(EnhancedVolcano)
library(TxDb.Rnorvegicus.UCSC.rn6.refGene)
library(topGO)
library(org.Rn.eg.db)
library(gridExtra)
library(DOSE)
library(clusterProfiler)
library(fgsea)
library(factoextra)
library(pathview)
library(enrichplot)
library(readr)
library(MuSiC)
library(MuSiC2)
library(ggpubr)
library(ggsignif)
library(rstatix)
library(scRNAseq)
library(VennDiagram)
library(stringr)
library(tibble)
```


#***Using DeSeq2***
We can now construct a DESeqDataSet object to formulate a starting point for our analysis. You need to add an appropriate design for analysis

#removing outliers identified to clean and rerun analysis
```{r}
se1<- se[,se$tissue_type == "cerebellum"] ##change tissue type here
#se1<-se1[ , !se1$bam_ID == c("CRBM_S32.bam","CRBM_S3.bam")] #remove specific outliers here
se1 <- se1[ rowSums(assay(se1)) >= 10, ]
#F_se<-se1[,se1$sex == "F"]
#M_se<-se1[,se1$sex == "M"]
#test<-M_se #[ , !se1$animal_ID ==] #remove specific outliers here
#colData(test)
#se1$group<-paste(se1$exposure, se1$batch_id, sep = "_") #create a group variable for easy comparison
#test$group
se1 <- se1[ , order( colnames( se1 ) ) ]
head(assay(se1))
#colData(se1)
#test$group<-factor(test$group, levels = c("Sham_M","Exp_M"))
dds <- DESeqDataSet(se1, 
                    design = ~ batch_id + exposure) #comparison done between high treatment and vehicle?
```

```{r}
dds$exposure<- factor(dds$exposure, levels = c("sham","low","medium","high"))

```


#***Exploratory analysis and visualization***
*transformations*
The variance stabilizing transformation (VST) a goal of stablizing the variance across the range of values. produce log2-like values for high counts. 
```{r}
colData(dds)

vsd <- vst(dds)
```


Now for visuals, plot a principal components analysis (PCA) using  ggplot2
```{r}
data <- plotPCA(vsd, 
                  intgroup = 'exposure' , 
                  returnData=TRUE)
data
percentVar <- round(100 * attr(data, 
                                 "percentVar"))
```
#PCA Plot
```{r}
tiff(file = 'Figures/PCA/Male_CRBLM_PCA.tiff', units="in", width=8, height=5, res=1000)
PCA <-ggplot(data, aes(PC1, 
                   PC2, 
                   color=exposure,
                   shape= exposure)) + 
  stat_ellipse(aes(color= exposure))+
  ggtitle("Male Cerebellum PCA")+
  geom_point(size=3) +
  scale_color_manual(values = c("#666666", "#FF66CC","#66CCCC","#9966FF")) +
  scale_shape_manual(values = c(19, 15,17,18)) +
  xlab(paste0("PC1: ",
              percentVar[1],
              "% variance")) +
  ylab(paste0("PC2: ",
              percentVar[2],
              "% variance"))+ 
  theme_bw() + 
  theme(panel.border = element_blank(), 
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black"))
PCA
dev.off()
PCA
```

#***Differential Expression***
Lets run a differential expression pipeline using the raw counts

This function will print out a message for the various steps it performs. These are described in more detail in the manual page for DESeq, which can be accessed by typing ?DESeq. Briefly these are: the estimation of size factors (controlling for differences in the sequencing depth of the samples), the estimation of dispersion values for each gene, and fitting a generalized linear model.
```{r}
dds <- DESeq(dds)
resultsNames(dds)
```
Calling results without any arguments will extract the estimated log2 fold changes and p values for the last variable in the design formula. If there are more than 2 levels for this variable, results will extract the results table for a comparison of the last level over the first level. Treatment 6 vs vehicle
```{r}
res_high<-results(dds, name = "exposure_high_vs_sham")
res_medium<-results(dds, name = "exposure_medium_vs_sham")
res_low<-results(dds, name = "exposure_low_vs_sham")

summary(res_high)
summary(res_medium)
summary(res_low)
```


#***Annotating results***
```{r}
columns(Rattus.norvegicus)
```

How to assign actual gene names to our counts. Using an annotation package for rattus. 
```{r}
res_high$symbol <- mapIds(Rattus.norvegicus, #which annotation package model to use
                     keys=row.names(res_high), #using rownames
                     column="SYMBOL", #add a column of symbols associated with row.names arguments for the model
                     keytype="GENEID", #our rownames are gene id's 
                     multiVals="first")
```

```{r}
res_high$TXname <- mapIds(Rattus.norvegicus, #which annotation package model to use
                     keys=row.names(res_high), #using rownames
                     column="TXNAME", #add a column of symbols associated with row.names arguments for the model
                     keytype="GENEID", #our rownames are gene id's 
                     multiVals="first")
```

```{r}
res_high$ONTOLOGY <- mapIds(Rattus.norvegicus,
                     keys=row.names(res_high),
                     column="ONTOLOGYALL",
                     keytype="GENEID",
                     multiVals="first")
```

```{r}
res_high$Genename <- mapIds(Rattus.norvegicus,
                     keys=row.names(res_high),
                     column="GENENAME", #now add a column for gene names or gene description
                     keytype="GENEID",
                     multiVals="first")
```

```{r}
res_high$GOID <- mapIds(Rattus.norvegicus,
                     keys=row.names(res_high),
                     column="GO", #now add a column for gene names or gene description
                     keytype="GENEID",
                     multiVals="first")
```

```{r}
res_high$ensembl <- mapIds(Rattus.norvegicus,
                     keys=row.names(res_high),
                     column="ENSEMBL", #now add a column for gene names or gene description
                     keytype="GENEID",
                     multiVals="first")
```
```{r}
res_high$ENTREZID <- mapIds(Rattus.norvegicus,
                     keys=row.names(res_high),
                     column="ENTREZID", #now add a column for gene names or gene description
                     keytype="GENEID",
                     multiVals="first")
```

```{r}
res_medium$symbol <- mapIds(Rattus.norvegicus, #which annotation package model to use
                     keys=row.names(res_medium), #using rownames
                     column="SYMBOL", #add a column of symbols associated with row.names arguments for the model
                     keytype="GENEID", #our rownames are gene id's 
                     multiVals="first")
```

```{r}
res_medium$TXname <- mapIds(Rattus.norvegicus, #which annotation package model to use
                     keys=row.names(res_medium), #using rownames
                     column="TXNAME", #add a column of symbols associated with row.names arguments for the model
                     keytype="GENEID", #our rownames are gene id's 
                     multiVals="first")

```

```{r}
res_medium$ONTOLOGY <- mapIds(Rattus.norvegicus,
                     keys=row.names(res_medium),
                     column="ONTOLOGYALL",
                     keytype="GENEID",
                     multiVals="first")
```

```{r}
res_medium$Genename <- mapIds(Rattus.norvegicus,
                     keys=row.names(res_medium),
                     column="GENENAME", #now add a column for gene names or gene description
                     keytype="GENEID",
                     multiVals="first")
```

```{r}
res_medium$GOID <- mapIds(Rattus.norvegicus,
                     keys=row.names(res_medium),
                     column="GO", #now add a column for gene names or gene description
                     keytype="GENEID",
                     multiVals="first")
```

```{r}
res_medium$ensembl <- mapIds(Rattus.norvegicus,
                     keys=row.names(res_medium),
                     column="ENSEMBL", #now add a column for gene names or gene description
                     keytype="GENEID",
                     multiVals="first")
```

```{r}
res_medium$ENTREZID <- mapIds(Rattus.norvegicus,
                     keys=row.names(res_medium),
                     column="ENTREZID", #now add a column for gene names or gene description
                     keytype="GENEID",
                     multiVals="first")
```

```{r}
res_low$symbol <- mapIds(Rattus.norvegicus, #which annotation package model to use
                     keys=row.names(res_low), #using rownames
                     column="SYMBOL", #add a column of symbols associated with row.names arguments for the model
                     keytype="GENEID", #our rownames are gene id's 
                     multiVals="first")

```

```{r}
res_low$TXname <- mapIds(Rattus.norvegicus, #which annotation package model to use
                     keys=row.names(res_low), #using rownames
                     column="TXNAME", #add a column of symbols associated with row.names arguments for the model
                     keytype="GENEID", #our rownames are gene id's 
                     multiVals="first")

 #take a look at the row names and the symbol column.
```

```{r}
res_low$ONTOLOGY <- mapIds(Rattus.norvegicus,
                     keys=row.names(res_low),
                     column="ONTOLOGYALL",
                     keytype="GENEID",
                     multiVals="first")

```

```{r}
res_low$Genename <- mapIds(Rattus.norvegicus,
                     keys=row.names(res_low),
                     column="GENENAME", #now add a column for gene names or gene description
                     keytype="GENEID",
                     multiVals="first")

```

```{r}
res_low$GOID <- mapIds(Rattus.norvegicus,
                     keys=row.names(res_low),
                     column="GO", #now add a column for gene names or gene description
                     keytype="GENEID",
                     multiVals="first")

```

```{r}
res_low$ensembl <- mapIds(Rattus.norvegicus,
                     keys=row.names(res_low),
                     column="ENSEMBL", #now add a column for gene names or gene description
                     keytype="GENEID",
                     multiVals="first")
```

```{r}
res_low$ENTREZID <- mapIds(Rattus.norvegicus,
                     keys=row.names(res_low),
                     column="ENTREZID", #now add a column for gene names or gene description
                     keytype="GENEID",
                     multiVals="first")
```


```{r}
High_resOrdered <- res_high[order(res_high$padj),]
medium_resOrdered <- res_medium[order(res_medium$padj),]
low_resOrdered <- res_low[order(res_low$padj),] #reorder the genes based on significance
```




#***Plotting results***
Heatmaps are a good way to visualize the most significant genes
```{r}
mat <- assay(vsd)[ head(order(res_high$padj),
                            20), #top 30 genes 
                       ]
#mat <- mat - rowMeans(mat)
df <- as.data.frame(colData(vsd)[,
                                     "exposure"])
rownames(df)<-colnames(mat)
colnames(df)[1]<- "Exposure"
```

```{r}
tiff(file = "Figures/heatmap/High_Male_CRBLM_heatmap.tiff", units="in", width=8, height=5, res=1000)
heatmap<-pheatmap(mat, 
         annotation_col=df,
         labels_row = High_resOrdered$symbol,
         fontsize = 6,
         scale = "row",
         show_colnames = F,
         cluster_cols = FALSE,
         main = "High Male Cerebellum Heatmap")
heatmap
dev.off()
heatmap
```

#***volcano plots***
```{r}
tiff(file = 'Figures/volcano/High_volcano_plot_CRBLM.tiff', units="in", width=8, height=5, res=1000)
M_vol1<- EnhancedVolcano(res_high,
    lab = res_high$symbol,
    x = 'log2FoldChange',
    y = 'pvalue',
    title = "Male Cerebellum",
    subtitle = "High Exposure ",
    ylim = c(0, -log10(10e-8)),
    xlim = c(log2(10e-3), -log2(10e-3)),
    FCcutoff = 1.0,
    pCutoff = 0.05,
    labSize = 3.0,
    colAlpha = 1,
    legendLabels=c('Not sig.','Log (base 2) FC','p-value',
      'p-value & Log (base 2) FC'),
    legendPosition = 'right',
    legendLabSize = 12,
    legendIconSize = 3.0,
    gridlines.major = FALSE,
    gridlines.minor = FALSE,
    border = 'full',
    borderWidth = 1.0,
    borderColour = 'black')
M_vol1
dev.off()

M_vol1
```
#Export Male TSV results

```{r}
High_resOrdered <- as.data.frame(High_resOrdered)
write_tsv(High_resOrdered, file="Result_spreadsheets/High_Male_CRBLM_results.tsv")
write.csv(High_resOrdered, file="Result_spreadsheets/High_Male_CRBLM_results.csv")
write.csv(as.data.frame(assay(se1)), file= "Result_spreadsheets/Male_CRBLM_counts.csv")
```


```{r}
mat <- assay(vsd)[ head(order(res_medium$padj),
                            20), #top 30 genes 
                       ]
#mat <- mat - rowMeans(mat)
df <- as.data.frame(colData(vsd)[,
                                     "exposure"])
rownames(df)<-colnames(mat)
colnames(df)[1]<- "Exposure"
```

```{r}
tiff(file = "Figures/heatmap/Medium_Male_CRBLM_heatmap.tiff", units="in", width=8, height=5, res=1000)
heatmap<-pheatmap(mat, 
         annotation_col=df,
         labels_row = medium_resOrdered$symbol,
         fontsize = 6,
         scale = "row",
         show_colnames = F,
         cluster_cols = FALSE,
         main = "Medium Male Cerebellum Heatmap")
heatmap
dev.off()
heatmap
```

```{r}
tiff(file = 'Figures/volcano/Medium_volcano_plot_CRBLM.tiff', units="in", width=8, height=5, res=1000)
M_vol1<- EnhancedVolcano(res_medium,
    lab = res_medium$symbol,
    x = 'log2FoldChange',
    y = 'pvalue',
    title = "Male Cerebellum",
    subtitle = "Medium Exposure ",
    ylim = c(0, -log10(10e-8)),
    xlim = c(log2(10e-3), -log2(10e-3)),
    FCcutoff = 1.0,
    pCutoff = 0.05,
    labSize = 3.0,
    colAlpha = 1,
    legendLabels=c('Not sig.','Log (base 2) FC','p-value',
      'p-value & Log (base 2) FC'),
    legendPosition = 'right',
    legendLabSize = 12,
    legendIconSize = 3.0,
    gridlines.major = FALSE,
    gridlines.minor = FALSE,
    border = 'full',
    borderWidth = 1.0,
    borderColour = 'black')
M_vol1
dev.off()

M_vol1
```

```{r}
medium_resOrdered <- as.data.frame(medium_resOrdered)
write_tsv(medium_resOrdered, file="Result_spreadsheets/Medium_Male_CRBLM_results.tsv")
write.csv(medium_resOrdered, file="Result_spreadsheets/Medium_Male_CRBLM_results.csv")
write.csv(as.data.frame(assay(se1)), file= "Result_spreadsheets/Medium_Male_CRBLM_counts.csv")
```

```{r}
mat <- assay(vsd)[ head(order(res_low$padj),
                            20), #top 30 genes 
                       ]
#mat <- mat - rowMeans(mat)
df <- as.data.frame(colData(vsd)[,
                                     "exposure"])
rownames(df)<-colnames(mat)
colnames(df)[1]<- "Exposure"
```

```{r}
tiff(file = "Figures/heatmap/low_Male_CRBLM_heatmap.tiff", units="in", width=8, height=5, res=1000)
heatmap<-pheatmap(mat, 
         annotation_col=df,
         labels_row = low_resOrdered$symbol,
         fontsize = 6,
         scale = "row",
         show_colnames = F,
         cluster_cols = FALSE,
         main = "Low Male Cerebellum Heatmap")
heatmap
dev.off()
heatmap
```

```{r}
tiff(file = 'Figures/volcano/Low_volcano_plot_CRBLM.tiff', units="in", width=8, height=5, res=1000)
M_vol1<- EnhancedVolcano(res_low,
    lab = res_low$symbol,
    x = 'log2FoldChange',
    y = 'pvalue',
    title = "Male Cerebellum",
    subtitle = "Low Exposure ",
    ylim = c(0, -log10(10e-8)),
    xlim = c(log2(10e-3), -log2(10e-3)),
    FCcutoff = 1.0,
    pCutoff = 0.05,
    labSize = 3.0,
    colAlpha = 1,
    legendLabels=c('Not sig.','Log (base 2) FC','p-value',
      'p-value & Log (base 2) FC'),
    legendPosition = 'right',
    legendLabSize = 12,
    legendIconSize = 3.0,
    gridlines.major = FALSE,
    gridlines.minor = FALSE,
    border = 'full',
    borderWidth = 1.0,
    borderColour = 'black')
M_vol1
dev.off()

M_vol1
```

```{r}
low_resOrdered <- as.data.frame(low_resOrdered)
write_tsv(low_resOrdered, file="Result_spreadsheets/Low_Male_CRBLM_results.tsv")
write.csv(low_resOrdered, file="Result_spreadsheets/Low_Male_CRBLM_results.csv")
write.csv(as.data.frame(assay(se1)), file= "Result_spreadsheets/Low_Male_CRBLM_counts.csv")
```


---------------

#**MALE gene enrichment analysis (GSEA)**
```{r}
x<- res_high[order(res_high$log2FoldChange, decreasing = TRUE),] #reorder the genes based on significance
head(x)
x<-na.omit(x)
gene_list<-x$log2FoldChange
names(gene_list)<-x$ensembl
head(gene_list)
```

```{r}
Mgse <- gseGO(geneList= gene_list, 
             ont ="ALL", 
             keyType = "ENSEMBL", 
             nPerm = 10000, 
             minGSSize = 3, 
             maxGSSize = 800, 
             pvalueCutoff = 0.05, 
             verbose = TRUE, 
             OrgDb = org.Rn.eg.db, 
             pAdjustMethod = "none")

```

```{r}
tiff(file = 'Figures/Enrichment/High_Male_CRBLM_gene_enrichment.tiff', units="in", width=7, height=5, res=1000)
dot<-dotplot(Mgse, showCategory=10, split=".sign") +
  facet_grid(.~.sign)+
  ggtitle("High Male Cerebellum Gene Enrichment")+ theme(axis.text.y = element_text(size = 5))    

dot
dev.off()
dot
```


# MALE KEGG pathview
```{r}
x<- res_high[order(res_high$log2FoldChange, decreasing = TRUE),]  #reorder the genes based on significance
x<-na.omit(x)
gene_list<-x$log2FoldChange
names(gene_list)<-x$ENTREZID
head(gene_list)
```


```{r}
kegg_organism = "rno"
Mkk2 <- gseKEGG(geneList     = gene_list,
               organism     = kegg_organism,
               nPerm        = 10000,
               minGSSize    = 3,
               maxGSSize    = 800,
               pvalueCutoff = 0.05,
               pAdjustMethod = "none",
               keyType       = "ncbi-geneid")
```

```{r}

tiff(file = 'Figures/Enrichment/High_Male_CRBLM_KEGG_enrichment.tiff', units="in", width=10, height=5, res=1000)
dot<-dotplot(Mkk2, showCategory=10, split=".sign") +
  facet_grid(.~.sign)+
  ggtitle("High Male Cerebellum KEGG Enrichment")+ theme(axis.text.y = element_text(size = 5))    

dot
dev.off()
dot
```

```{r}
# Produce the native KEGG plot (PNG)
#male_cerebellum <- pathview(gene.data=gene_list, pathway.id="05012", species = kegg_organism)

```

```{r}
x<- res_medium[order(res_medium$log2FoldChange, decreasing = TRUE),] #reorder the genes based on significance
head(x)
x<-na.omit(x)
gene_list<-x$log2FoldChange
names(gene_list)<-x$ensembl
head(gene_list)
```

```{r}
Mgse <- gseGO(geneList= gene_list, 
             ont ="ALL", 
             keyType = "ENSEMBL", 
             nPerm = 10000, 
             minGSSize = 3, 
             maxGSSize = 800, 
             pvalueCutoff = 0.05, 
             verbose = TRUE, 
             OrgDb = org.Rn.eg.db, 
             pAdjustMethod = "none")

```

```{r}
tiff(file = 'Figures/Enrichment/Medium_Male_CRBLM_gene_enrichment.tiff', units="in", width=7, height=5, res=1000)
dot<-dotplot(Mgse, showCategory=10, split=".sign") +
  facet_grid(.~.sign)+
  ggtitle("Medium Male Cerebellum Gene Enrichment")+ theme(axis.text.y = element_text(size = 5))    

dot
dev.off()
dot
```


```{r}
x<- res_medium[order(res_medium$log2FoldChange, decreasing = TRUE),]  #reorder the genes based on significance
x<-na.omit(x)
gene_list<-x$log2FoldChange
names(gene_list)<-x$ENTREZID
head(gene_list)
```


```{r}
kegg_organism = "rno"
Mkk2 <- gseKEGG(geneList     = gene_list,
               organism     = kegg_organism,
               nPerm        = 10000,
               minGSSize    = 3,
               maxGSSize    = 800,
               pvalueCutoff = 0.05,
               pAdjustMethod = "none",
               keyType       = "ncbi-geneid")
```

```{r}

tiff(file = 'Figures/Enrichment/Medium_Male_CRBLM_KEGG_enrichment.tiff', units="in", width=10, height=5, res=1000)
dot<-dotplot(Mkk2, showCategory=10, split=".sign") +
  facet_grid(.~.sign)+
  ggtitle("Medium Male Cerebellum KEGG Enrichment")+ theme(axis.text.y = element_text(size = 5))    

dot
dev.off()
dot
```

```{r}
x<- res_low[order(res_low$log2FoldChange, decreasing = TRUE),] #reorder the genes based on significance
head(x)
x<-na.omit(x)
gene_list<-x$log2FoldChange
names(gene_list)<-x$ensembl
head(gene_list)
```

```{r}
Mgse <- gseGO(geneList= gene_list, 
             ont ="ALL", 
             keyType = "ENSEMBL", 
             nPerm = 10000, 
             minGSSize = 3, 
             maxGSSize = 800, 
             pvalueCutoff = 0.05, 
             verbose = TRUE, 
             OrgDb = org.Rn.eg.db, 
             pAdjustMethod = "none")

```

```{r}
tiff(file = 'Figures/Enrichment/Low_Male_CRBLM_gene_enrichment.tiff', units="in", width=7, height=5, res=1000)
dot<-dotplot(Mgse, showCategory=10, split=".sign") +
  facet_grid(.~.sign)+
  ggtitle("Low Male Cerebellum Gene Enrichment")+ theme(axis.text.y = element_text(size = 5))    

dot
dev.off()
dot
```


```{r}
x<- res_low[order(res_low$log2FoldChange, decreasing = TRUE),]  #reorder the genes based on significance
x<-na.omit(x)
gene_list<-x$log2FoldChange
names(gene_list)<-x$ENTREZID
head(gene_list)
```


```{r}
kegg_organism = "rno"
Mkk2 <- gseKEGG(geneList     = gene_list,
               organism     = kegg_organism,
               nPerm        = 10000,
               minGSSize    = 3,
               maxGSSize    = 800,
               pvalueCutoff = 0.05,
               pAdjustMethod = "none",
               keyType       = "ncbi-geneid")
```

```{r}

tiff(file = 'Figures/Enrichment/Low_Male_CRBLM_KEGG_enrichment.tiff', units="in", width=10, height=5, res=1000)
dot<-dotplot(Mkk2, showCategory=10, split=".sign") +
  facet_grid(.~.sign)+
  ggtitle("Low Male Cerebellum KEGG Enrichment")+ theme(axis.text.y = element_text(size = 5))    

dot
dev.off()
dot
```

#plotting individual genes of interest
Using normalized counts (prebuilt into DESEQ2) which normalizes counts by the estimated size factors (or normalization factors if these were used) and adds a pseudocount of 1/2 to allow for log scale plotting.
**drug metabolizing enzymes**
```{r}
top_gene<-read.csv("gene_list_drug_metabolizing_kegg_CRBLM.csv", header = FALSE)
top_gene<-as.character(top_gene$V1)

setdiff(top_gene,rownames(counts(dds))) #check if all genes of interest can be found in the rownames of the deseq object --> if not you will not be able to run the block of code. Character(0) or integer(0) is a what you want to see
```
```{r}
stat.test<-read.csv("CRBLM_FDR.csv")
stat.test$symbol<-factor(stat.test$symbol)
stat.test$group2<-as.factor(stat.test$group2)
stat.test$group1<-as.factor(stat.test$group1)
stat.test
```

```{r}
data_1 <- split(as.tibble(stat.test[,2:4]), as.factor(stat.test$symbol))
```


```{r message=FALSE}
plot_list<-list()
for (i in unique(1:length(top_gene))){
  gene <- top_gene[i]
  b<- plotCounts(dds, gene = gene, intgroup = "exposure",normalized = TRUE, returnData = TRUE)
  z = max(b$count) + (max(b$count)*0.05)
  d <- ggplot(b, aes(factor(exposure), count))+
    geom_violin(mapping = aes(x = exposure, 
                              y = count, 
                              fill = exposure))+
    #adding jitter to avoid overplotting
    geom_point(mapping = aes(x = exposure, 
                             y = count, 
                             fill = exposure, 
                             shape= exposure), 
               size = 5, position = position_jitter(width = 0.3, height=0))+ 
    scale_y_continuous(expand = c(0,0) , limits = c(0,z+z*0.4)) +
    stat_summary(mapping = aes(x = exposure, y = count), 
                 geom = 'crossbar', 
                 fun= 'mean', 
                 colour = 'black', 
                 width=0.2)+
    scale_fill_manual(values = c("#666666", "#FF66CC","#66CCCC","#9966FF")) +
    scale_shape_manual(values = c(19, 15,17,18)) +
  labs(x = NULL, y = "Normalized counts")+ # changes the y axis label, removes x axis label
  theme_classic(base_size = 20)+# changes the overall style of the plot
  ggtitle(paste0(res_low$symbol[gene])) +
  theme(
    axis.text.x = element_text(colour = "black"),
    axis.text.y = element_text(colour = "black"))+
  stat_pvalue_manual(data_1[[i]], 
    y.position = z, step.increase = 0.2,
    label = "p = {scales::pvalue(p.adj)}", hide.ns = TRUE
    )
  plot_list[[gene]] <- d
}

head(plot_list)
```


```{r warning=FALSE}
# Export into pdf: display multiple plots on the same page
ggexport(
  plotlist = plot_list, filename = "Figures/male_CRBLM_cyp_plot.pdf", 
  ncol = 1, nrow = 1, height = 7, width = 7, res = 600,pointsize = 8
)
```

